---
title: "Assessment 2 for SRS"
author: "Team 7"
date: "2025-02-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Try**

```{r load data}
# load the data
setwd("~/Desktop/SRS/Assessment2")
ourdata <- read.csv("finalbis.csv", header = TRUE)
head(ourdata)
```

```{r dist of added value}
# See the distribution of added value
y <- ourdata$added_value
hist(y,main="Distribution of Added Value", xlab="Added Value", col="grey", probability = TRUE)
lines(density(y), col = "red", lwd = 2)
print(c(mean(y), sd(y)))
```

Clean the data, delete all NAs and unrelated rows:
```{r Data clean}
#install.packages("dplyr")
library(dplyr)
# Delete unrelated column
newdata <- ourdata[, -c(1,2,4,18,19,20,21,22,23,24,25,26,27)]
# Delete Na rows
newdata$continuation[newdata$continuation == "n/a"] <- NA
newdata <- newdata[!is.na(newdata$satisfied_teaching) & !is.na(newdata$continuation), ]
newdata$continuation <- as.numeric(newdata$continuation)
newdata$INSTITUTION_CODE <- as.factor(newdata$INSTITUTION_CODE)

newdata <- newdata %>%
  mutate(
    ethnic_group = case_when(
      White.ethnic.group == pmax(White.ethnic.group, Black.ethnic.group, Asian.ethnic.group, Mixed.ethnic.group, Other.ethnic.group) ~ "White",
      Black.ethnic.group == pmax(White.ethnic.group, Black.ethnic.group, Asian.ethnic.group, Mixed.ethnic.group, Other.ethnic.group) ~ "Black",
      Asian.ethnic.group == pmax(White.ethnic.group, Black.ethnic.group, Asian.ethnic.group, Mixed.ethnic.group, Other.ethnic.group) ~ "Asian",
      Mixed.ethnic.group == pmax(White.ethnic.group, Black.ethnic.group, Asian.ethnic.group, Mixed.ethnic.group, Other.ethnic.group) ~ "Mixed",
      Other.ethnic.group == pmax(White.ethnic.group, Black.ethnic.group, Asian.ethnic.group, Mixed.ethnic.group, Other.ethnic.group) ~ "Other",
      TRUE ~ "Unknown"
    )
  )
newdata <- newdata %>%
  mutate(
    sex = case_when(
      Men == pmax(Men, Women) ~ "Men",
      Women == pmax(Men, Women) ~ "Women",
      TRUE ~ "Unknown"
    )
  )
newdata$ethnic_group <- as.factor(newdata$ethnic_group)

newdata <- newdata %>%
  select(-White.ethnic.group, -Black.ethnic.group, -Asian.ethnic.group, -Mixed.ethnic.group, -Other.ethnic.group, -Men, -Women)
```

Try to fit the linear modal for all variable:
```{r Linear Model}
names(newdata)
```

```{r}
library(car)
vif(model1)
```
```{r}
new_model <- lm(career_after_15_month ~ . - INSTITUTION_CODE, data = newdata)
vif(new_model)
step(new_model)
```


```{r}
final.model <- lm(formula = career_after_15_month ~ satisfied_teaching + satisfied_feedback + 
    students_staff_ratio + avg_entry_tariff + added_value + continuation + 
    Total, data = newdata)
summary(final.model)
plot(final.model)
```

Bayesian:
```{r}
library(rjags)
```

```{r}
model_string <- "
model {
  # Prior for the precision
  tau~dgamma(a, b)
  
  # Compute the variance
  sigma2 = 1/tau

  # Prior for beta
  tau0 = 1/sigma02
  for(j in 1:8){
  beta[j]~dnorm(mu0,tau0)
  }
  
  # Likelihood
  for (i in 1:N) {
    career[i] ~ dnorm(mu[i], tau)
    mu[i] <- beta[1] + beta[2] * teaching[i] + beta[3] * feedback[i] +
             beta[4] * staff_ratio[i] + beta[5] * entry[i] +
             beta[6] * added[i] + beta[7] * continuation[i] + beta[8] * total[i]
  }
  
}
"
```

```{r}
# hyperparameter for the betas and tau
mu0 = 0; sigma02 = 1000; a = 0.1; b = 0.1

data_jags <- list(
  N = nrow(newdata),
  career = newdata$career_after_15_month,
  teaching = as.vector(scale(newdata$satisfied_teaching)),
  feedback = as.vector(scale(newdata$satisfied_feedback)),
  staff_ratio = as.vector(scale(newdata$students_staff_ratio)),
  entry = as.vector(scale(newdata$avg_entry_tariff)),
  added = as.vector(scale(newdata$added_value)),
  continuation = as.vector(scale(newdata$continuation)),
  total = as.vector(scale(newdata$Total)),
  mu0 = mu0,
  sigma02 = sigma02,
  a = a,
  b = b
)
```

```{r}
# compile the model
model.jags = jags.model(textConnection(model_string), n.chains = 3, data = data_jags)
```
```{r}
update(model.jags, 10000, progress.bar = "none")
res = coda.samples(model.jags, variable.names = c("beta", "sigma2"), n.iter = 50000, 
                   thin = 100, progress.bar = "none")
summary(res)
```

Plot the results to check for convergence
Check for convergence before looking at posterior dist'n summaries
```{r}
par(mfrow = c(2, 4))
for (i in 1:8) {
  param_name <- paste0("beta[", i, "]")
  plot(res[, param_name], main = param_name)
}
```
Evaluate autocorrelation functions
```{r}
par(mfrow=c(1,3))
acf(res[[1]][,"beta[1]"],lag.max=100)
acf(res[[1]][,"beta[2]"],lag.max=100)
acf(res[[1]][,"sigma2"],lag.max=100)
```

Compute the Gelman-Rubin statistic
```{r}
gelman.diag(res)
```

Now, consider the robustify the regression, the effects of extreme observations or “outliers” on regressing results can be diminished by using a t
 distribution for the observations. 
 
```{r}
model_string_robust <- "
model {
  # Prior for the precision
  tau~dgamma(a, b)
  
  # Compute the variance
  sigma2 = 1/tau

  # Prior for beta
  tau0 = 1/sigma02
  for(j in 1:8){
  beta[j]~dnorm(mu0,tau0)
  }
  
  # Likelihood
  for (i in 1:N) {
    career[i] ~ dt(mu[i], tau, df)
    mu[i] <- beta[1] + beta[2] * teaching[i] + beta[3] * feedback[i] +
             beta[4] * staff_ratio[i] + beta[5] * entry[i] +
             beta[6] * added[i] + beta[7] * continuation[i] + beta[8] * total[i]
  }
  
}
"

# hyperparameter for the betas and tau
mu0 = 0; sigma02 = 1000; a = 0.1; b = 0.1; df = 9

data_jags_robust <- list(
  N = nrow(newdata),
  career = newdata$career_after_15_month,
  teaching = as.vector(scale(newdata$satisfied_teaching)),
  feedback = as.vector(scale(newdata$satisfied_feedback)),
  staff_ratio = as.vector(scale(newdata$students_staff_ratio)),
  entry = as.vector(scale(newdata$avg_entry_tariff)),
  added = as.vector(scale(newdata$added_value)),
  continuation = as.vector(scale(newdata$continuation)),
  total = as.vector(scale(newdata$Total)),
  mu0 = mu0,
  sigma02 = sigma02,
  a = a,
  b = b,
  df = df
)

model.robust <- jags.model(textConnection(model_string_robust), n.chains = 3, data = data_jags_robust)
```
 
```{r}
update(model.robust, n.iter = 10000)
res.sample.robust <- coda.samples(model.robust, variable.names = c("beta", "sigma2"), n.iter = 50000,
                   thin = 100, progress.bar = "none")
```

```{r}
par(mfrow = c(2, 4))
for (i in 1:8) {
  param_name <- paste0("beta[", i, "]")
  plot(res.sample.robust[, param_name], main = param_name)
}
```

```{r}
summary(res.sample.robust)
```


















